# MME: No sexID strat


# Het, gay/lesbian, bi population, sexual activity stratification only

In this model, we do not stratify by sexual identity or restrict to heterosexual sexual identity. The analysis procedure is exactly the same as for the heterosexual-only population.

## Estimate the number of new partnerships with each sex

```{r}
het_partners <- all_sex %>%
    group_by(r_sex, r_sexact) %>%
    summarise(partners = weighted.mean(hetnonew, w = total_wt)) %>%
    mutate(rp_sex = ifelse(r_sex == "m", 'w', 'm')) %>%
    rowwise() %>%
    mutate(prop = get_proportion(all_sex,
                                 q_sex = r_sex,
                                 q_sexact = r_sexact),
           n_partners = partners * prop) %>%
    ungroup()

sam_partners <- all_sex %>%
    group_by(r_sex, r_sexact) %>%
    summarise(partners = weighted.mean(samnonew, w = total_wt)) %>%
    mutate(rp_sex = ifelse(r_sex == "m", 'm', 'w')) %>%
    rowwise() %>%
    mutate(prop = get_proportion(all_sex,
                                 q_sex = r_sex,
                                 q_sexact = r_sexact),
           n_partners = partners * prop) %>%
    ungroup()

sexid_nostrat_rep <- rbind(het_partners, sam_partners) %>%
  arrange(r_sex, r_sexact)
  
format_table(sexid_nostrat_rep)
```

Note that the sum of the `prop` column is now 2:

```{r}
sum(sexid_nostrat_rep$prop)
```

That's because each demographic group has two entries in the table: one for partnerships with the same sex and one for the opposite sex. This format makes it easier to do the balancing, as above.

Next, let's distribute the partnerships across activity levels.

## Distribute partnerships

```{r}
sexid_nostrat_offered <-  sexid_nostrat_rep %>%
  group_by(rp_sex, r_sex) %>%
  mutate(prop_to_r = n_partners / sum(n_partners)) %>%
  select(rp_sex,
         r_sex,
         rp_sexact = r_sexact,
         prop_to_r) %>%
  ungroup() %>%
  arrange(rp_sex, r_sex)
format_table(sexid_nostrat_offered)
```

Here, the percentages refer to the proportion of partnerships with people of sex `rp_sex` that are offered to `r_sex` by people with sex `rp_sex` and activity `rp_sexact`. For example, high-activity men offer about 91% percent of male partnerships to men, and high-activity women offer about 66% of female partnerships to men.

Now, we join the offered table and the reported table.

```{r}
sexid_nostrat_joined <- left_join(sexid_nostrat_offered, sexid_nostrat_rep, by = c("r_sex", "rp_sex"))
format_table(sexid_nostrat_joined)
```

The multiplication gives us the distributed parterships. Then, we define a single variable that describes the two demographic variables (sex and SA group):

```{r}
partner_dist_sexid_nostrat <- sexid_nostrat_joined %>%
  mutate(d_partners = partners * prop_to_r,
         r_demo = paste(r_sex, r_sexact, sep="_"),
         rp_demo = paste(rp_sex, rp_sexact, sep="_")) %>%
  select(r_demo, rp_demo, d_partners, prop)
format_table(partner_dist_sexid_nostrat)
```

So, men with high sexual activity reported `r format(sexid_nostrat_rep$partners[1], digits = 3)` partners with women. In the distributed table, these partners are distributed across women with high sexual activity and women with low sexual activity.

```{r}
men_high_sa <- partner_dist_sexid_nostrat %>%
  filter(r_demo == "m_high" & (rp_demo == "w_high" | rp_demo == "w_low"))
format_table(men_high_sa)
sum(men_high_sa$d_partners)
```

Similarly, high-SA men reported `r format(sexid_nostrat_rep$partners[2], digits = 3)` with men on average, which are distributed across men by sexual activity:

```{r}
men_high_sa <- partner_dist_sexid_nostrat %>%
  filter(r_demo == "m_high" & (rp_demo == "m_high" | rp_demo == "m_low"))
format_table(men_high_sa)
sum(men_high_sa$d_partners)
```

## Balancing

Join the distributed data frame with itself:

```{r}
sexid_nostrat_bidi <- partner_dist_sexid_nostrat %>%
    # do a self-join t calculate partners from perspective of rp
    inner_join(partner_dist_sexid_nostrat,
               by = c("r_demo" = "rp_demo",
                      "rp_demo" = "r_demo"),
               suffix = c('.r', '.rp'))
sexid_nostrat_bidi %>%
  format_table()
```


```{r}
theta <- 0.5
b_sexid_nostrat_bidi <- sexid_nostrat_bidi %>%
    mutate(np_r = d_partners.r * prop.r,
           np_rp = d_partners.rp * prop.rp) %>%
    mutate(imbalance = np_r / np_rp,
           corrected_r = d_partners.r / imbalance^(1 - theta),
           corrected_rp = d_partners.rp * imbalance^theta,
           cnr = corrected_r * prop.r,
           cnrp = corrected_rp * prop.rp) %>%
    select(r_demo,
           rp_demo,
           prop.r,
           np_r,
           prop.rp,
           np_rp,
           d_partners.r,
           corrected_r,
           d_partners.rp,
           corrected_rp,
           imbalance,
           cnr,
           cnrp) %>%
  arrange(r_demo, rp_demo)

format_table(b_sexid_nostrat_bidi)
all(with(b_sexid_nostrat_bidi, abs(cnr - cnrp) < .Machine$double.eps))
```

Balanced!

### Compare

```{r}
library(stringr)
sexid_nos_by_grp <- b_sexid_nostrat_bidi %>%
  mutate(rp_sex = str_split(rp_demo, "_", simplify = TRUE)[, 1]) %>%
  group_by(r_demo, rp_sex) %>%
  summarize(bal_p_pc = sum(corrected_r))

sexid_nos_compare <- sexid_nostrat_rep %>%
  mutate(r_demo = paste(r_sex, r_sexact, sep = "_")) %>%
  select(r_demo, rp_sex, orig_p_pc = partners) %>%
  left_join(sexid_nos_by_grp, by = c("r_demo", "rp_sex"))
format_table(sexid_nos_compare)
```

```{r}
# plot with nas
library(ggplot2)
plot_mixing_matrix(b_sexid_nostrat_bidi)
```


