# Goal of Analysis

The goal is to estimate the annual number of new partnerships between each group of people in the model.

In the paper, we use three model populations:

1. Heterosexual population, sexual activity stratification only
2. Het, gay/lesbian, bi population, sexual activity stratification only
3. Het, gay/lesbian, bi population, sexual activity and sexID stratification

In the Natsal data, we know the sex, self-reported sexual identity (sexID), and sexual activity group (SA) of the respondents, as well as the sex of the respondents' partners. The main problem is that the respondents don't report (and/or don't know) the sexID or SA of their partner. So, we know the total number of partners that each group has with each sex, but we'll have to distribute these partnerships among sexIDs and SAs.

Because there doesn't seem to be much information on how people mix regarding sexIDs or SAs, we make the proportionate assumption, which amounts to assuming that people have no preference for sexID or SA - they simply choose randomly (uniformly) among what is available to them.

The analysis has three steps:

1. Estimate the number of new partnerships with each sex
2. Distribute partnerships across sexual identities and sexual activity groups using the proportionality assumption
3. Balance the resulting distributed partnerships

We carry out the analysis for each model population.

## Data prep

First, we read in the Natsal data and prepare it for analysis. We use the following R packages:

```{r message=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
library(kableExtra)
```

This is a function to pretty-print tables.
```{r}
format_table <- function(df) {
  df %>%
    knitr::kable(booktabs = TRUE, digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = "striped",
                              latex_options = c("scale_down", "striped"))
}
```

## Relevant Variables in Natsal

The following variables in the Natsal data set are relevant to our analysis:

* rsex: respondent's sex
    - 1 is male
    - 2 is female

* sexid: sexual identity
    - 1: heterosexual / straight
    - 2: gay / lesbian
    - 3: bisexual
    - 4: other
    - 9: not answered

* everhet: ever had a heterosexual sexual partnership
    - 1 is yes
    - 0 is no
    - 9 is "unclassifiable"

* eversam: Ever had a same-sex partnership
    - 1 is yes
    - 2 is no
    - 9: NA
    - -1: not answered

* hetnonew: total number of new heterosexual partners in the past year
    - 0: if het1yr=0 | (het1yr=1 & hetnewp=2)
    - 1: if het1yr=1 & hetnewp=1
    - 99: not answered if missing
    - -1: not applicable if het1yr=-1

* samnonew: no. of new hom. sex partners, last year
    - 0 if sam1yr=0 | (sam1yr=1 & samnewp=2)
    - 1 if sam1yr=1 & samnewp=1
    - copy hnonewp if sam1yr>1 & sam1yr<995
    - 999: not answered if missing
    - -1: not applicable if sam1yr=-1

* totnewyr: no. of new het. & hom. sex partners, last year
    - compute hetnonew + samnonew if hetnonew>=0 & hetnonew<999 & samnonew>=0 & samnonew<999
    - else hetnonew if samnonew=-1 & hetnonew>=0 & hetnonew<999
    - else samnonew if hetnonew=-1 & samnonew>=0 & samnonew<999
    - 999: not answered if hetnonew=999 | samnonew=999
    - -1: not applicable if hetnonew =-1 | samnonew =-1



## Analysis

Load NATSAL data in R form:
```{r}
load("../../old/sfceSO/natsal_R_df.rda")
```

We focus on a higher-risk age group: 18-year-olds to 35-year-olds. 
```{r}
MIN_AGE <- 18
MAX_AGE <- 35

natsal_hr <- filter(natsal_R_df, dage >= MIN_AGE & dage <= MAX_AGE)
```

Next, we make new indicator variables, recode some variables to make their values more transparent, and select only relevant variables.

We make two other choices:

1. Exclude anyone who doesn't report a sexual identity of heterosexual, gay, or bisexual. A large percentage (over 99%) report one of these three sexual identities.
```{r}
natsal_filt <- natsal_hr %>%
  filter(sexid == 1 | sexid == 2 | sexid == 3)
nrow(natsal_filt) / nrow(natsal_hr)
```

2. Remove missing partner data. For `hetnonew`, "99" indicates missingness and "-1" indicates "not applicable". For `samnonew`, "999" indicates missingness and "-1" is not applicable. Again, the missing data is a small proportion of the total (less than 3%).
```{r}
natsal_filt2 <- natsal_filt %>%
  filter(hetnonew < 99 & samnonew < 995 & hetnonew >= 0 & samnonew >= 0)
nrow(natsal_filt2)/nrow(natsal_filt)
```

Now, we recode the sexid to allow for easier interpretation.

```{r}
recode_sexid <- function(sexid){
  new_sexid <- rep(NA, length(sexid))
  new_sexid[sexid == 1] <- 'het'
  new_sexid[sexid == 2] <- 'gay'
  new_sexid[sexid == 3] <- 'bi'
  return(new_sexid)
}

all_sex <- natsal_filt2 %>%
  # make new indicator for female, eversam and everhet
  mutate(
      eversam_ind = (eversam == 1),
      everhet_ind = (everhet == 1),
      r_sex = ifelse(rsex == 1, "m", "w"),
      r_sexid = recode_sexid(sexid)
  ) %>%
  select(r_sex, r_sexid, eversam_ind, everhet_ind, hetnonew, samnonew, totnewyr, total_wt)
```

In all, we end up with `r nrow(all_sex)` observations, which is `r format(nrow(all_sex)/nrow(natsal_hr)*100, digits = 3)`% of the original respondents between `r MIN_AGE` and `r MAX_AGE` years old.

Next, we define the 'sexual activity group', based on whether someone reports having 0-1 (low risk) or 2+ (high risk) new sex partners in the past year. 

```{r}
def_sex_act_group <- function(totnewyr){
  lent <- length(totnewyr)
  ret <- rep(NA, lent)
  ret[totnewyr == 0 | totnewyr == 1] <- 'low'
  ret[totnewyr >= 2] <- 'high'
  return(ret)
}

all_sex$r_sexact <- def_sex_act_group(all_sex$totnewyr)
```

These are the first few rows of the final cleaned data:
```{r asis=TRUE}
format_table(head(all_sex))
```
