---
output:
  html_document: default
  word_document: default
---
# Goal of Analysis

The goal is to estimate the annual number of new partnerships between each group of people in the model.

In the paper, we use three model populations:

1. Heterosexual population, sexual activity stratification only
2. Het, gay/lesbian, bi population, sexual activity stratification only
3. Het, gay/lesbian, bi population, sexual activity and sexID stratification

In the Natsal data, we know the sex, self-reported sexual identity (sexID), and sexual activity group (SA) of the respondents, as well as the sex of the respondents' partners. The main problem is that the respondents don't report (and/or don't know) the sexID or SA of their partner. So, we know the total number of partners that each group has with each sex, but we'll have to distribute these partnerships among sexIDs and SAs.

Because there doesn't seem to be much information on how people mix regarding sexIDs or SAs, we make the proportionate assumption, which amounts to assuming that people have no preference for sexID or SA - they simply choose randomly (uniformly) among what is available to them.

The analysis has three steps:

1. Estimate the number of new partnerships with each sex
2. Distribute partnerships across sexual identities and sexual activity groups using the proportionality assumption
3. Balance the resulting distributed partnerships

We carry out the analysis for each model population.

## Data prep

First, we read in the Natsal data and prepare it for analysis. We use the following R packages:

```{r message=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
library(kableExtra)
library(stringr)
library(usethis)
```

This is a function to pretty-print tables.
```{r}
format_table <- function(df) {
  df %>%
    knitr::kable(booktabs = TRUE, digits = 3) %>%
    kableExtra::kable_styling(bootstrap_options = "striped",
                              latex_options = c("scale_down", "striped"))
}
```


## Relevant Variables in Natsal

The following variables in the Natsal data set are relevant to our analysis:

* rsex: respondent's sex
    - 1 is male
    - 2 is female

* sexid: sexual identity
    - 1: heterosexual / straight
    - 2: gay / lesbian
    - 3: bisexual
    - 4: other
    - 9: not answered

* hetnonew: total number of new heterosexual partners in the past year
    - 0: if het1yr=0 | (het1yr=1 & hetnewp=2)
    - 1: if het1yr=1 & hetnewp=1
    - 99: not answered if missing
    - -1: not applicable if het1yr=-1

* samnonew: no. of new hom. sex partners, last year
    - 0 if sam1yr=0 | (sam1yr=1 & samnewp=2)
    - 1 if sam1yr=1 & samnewp=1
    - copy hnonewp if sam1yr>1 & sam1yr<995
    - 999: not answered if missing
    - -1: not applicable if sam1yr=-1

* totnewyr: no. of new het. & hom. sex partners, last year
    - compute hetnonew + samnonew if hetnonew>=0 & hetnonew<999 & samnonew>=0 & samnonew<999
    - else hetnonew if samnonew=-1 & hetnonew>=0 & hetnonew<999
    - else samnonew if hetnonew=-1 & samnonew>=0 & samnonew<999
    - 999: not answered if hetnonew=999 | samnonew=999
    - -1: not applicable if hetnonew =-1 | samnonew =-1



## Analysis

Load NATSAL data in R form:
```{r}
load("../../natsal_R_df.rda")
```

All the subsetting and analyses are done using functions the `survey` package. To do this, we first have to specify the survey design:

```{r}
library(survey)
natsal_des <- svydesign(data = natsal_R_df, ids=~psu_scrm,
                        strata=~stratagrp2, weights=~total_wt, nest = TRUE)
```

We focus on a higher-risk age group: 18-year-olds to 34-year-olds.

```{r}
MIN_AGE <- 18
MAX_AGE <- 34

natsal_hr <- subset(natsal_des, dage >= MIN_AGE & dage <= MAX_AGE)
```

### Re-code missing data

```{r}
natsal_hr <- update(natsal_hr,
                    hetnonew = ifelse(hetnonew < 99 & hetnonew >= 0, hetnonew, NA),
                    samnonew = ifelse(samnonew < 995 & samnonew >= 0, samnonew, NA),
                    totnewyr = ifelse(totnewyr < 999 & totnewyr >= 0, totnewyr, NA),
                    sexid = ifelse(sexid == 9, NA, sexid))
```

## Descriptive Analyses

### Sample Size

```{r}
svytotal(~total_wt, natsal_hr)
```

### Age

```{r}
svymean(~dage, natsal_hr)
```

### Sex Distribution

```{r}
svytotal(~factor(rsex), natsal_hr)
svymean(~factor(rsex), natsal_hr)
```

### Sexual Identity

```{r}
svytotal(~factor(sexid, exclude = ""), natsal_hr)
svymean(~factor(sexid, exclude = ""), natsal_hr)
```

### Number of new heterosexual partners

```{r}
svytotal(~factor(hetnonew >= 2, exclude = ""), natsal_hr)
svymean(~factor(hetnonew >= 2, exclude = ""), natsal_hr)
```

### Number of new same-sex partners in the last year

```{r}
svytotal(~factor(samnonew >= 2, exclude = ""), natsal_hr)
svymean(~factor(samnonew >= 2, exclude = ""), natsal_hr)
```

### Total number of new partners in the last year

```{r}
svytotal(~factor(totnewyr >= 2, exclude = ""), natsal_hr)
svymean(~factor(totnewyr >= 2, exclude = ""), natsal_hr)
```


## Further data preparation

Next, we make new indicator variables, recode some variables to make their values more transparent, and select only relevant variables.

We make two other choices:

1. Exclude anyone who doesn't report a sexual identity of heterosexual, gay, or bisexual. A large percentage (over 99%) report one of these three sexual identities.
```{r}
natsal_sexid <- subset(natsal_hr, sexid == 1 | sexid == 2 | sexid == 3)
nrow(natsal_sexid) / nrow(natsal_hr)
```

2. Now, we impute missing partnership numbers.

```{r}
library(HotDeckImputation)
vars <- data.matrix(
  natsal_sexid$variables[, c("dage", "sexid", "hetnonew", "samnonew")]
)
imputed <- data.frame(impute.CPS_SEQ_HD(vars, 
                                    covariates = c(1, 2),
                                    initialvalues = 0,
                                    modifyinplace = FALSE))
natsal_imp <- update(natsal_sexid,
                     hetnonew = imputed$hetnonew,
                     samnonew = imputed$samnonew)
```

Next, we update the total number of partners in a new year as a sum of the two new variables.

```{r}
natsal_imp <- update(natsal_imp,
                     totnewyr = hetnonew + samnonew)
```

Now, we recode the sexid to allow for easier interpretation.

```{r}
recode_sexid <- function(sexid){
  new_sexid <- rep(NA, length(sexid))
  new_sexid[sexid == 1] <- 'het'
  new_sexid[sexid == 2] <- 'gay'
  new_sexid[sexid == 3] <- 'bi'
  return(new_sexid)
}

natsal_recoded_sexid <- update(natsal_imp,
                               r_sexid = recode_sexid(sexid),
                               eversam_ind = (eversam == 1),
                               everhet_ind = (everhet == 1),
                               r_sex = ifelse(rsex == 1, "m", "w"))
```

Next, we define the 'sexual activity group', based on whether someone reports having 0-1 (low risk) or 2+ (high risk) new sex partners in the past year. 

```{r}
def_sex_act_group <- function(totnewyr){
  # total partnerships
  lent <- length(totnewyr)
  ret <- rep(NA, lent)
  ret[totnewyr == 0 | totnewyr == 1] <- 'low'
  ret[totnewyr >= 2] <- 'high'
  return(ret)
}

def_sex_act_group_het <- function(hetnonew){
  # total partnerships
  lent <- length(hetnonew)
  ret <- rep(NA, lent)
  ret[hetnonew == 0 | hetnonew == 1] <- 'low'
  ret[hetnonew >= 2] <- 'high'
  return(ret)
}

natsal_sag <- update(natsal_recoded_sexid,
                     r_sexact = def_sex_act_group(totnewyr),
                     r_sexact_het = def_sex_act_group_het(hetnonew))
```

## Redefine partnerships

Create new variables `m_nonew` and `w_nonew`. This is the final natsal dataframe for analysis.

```{r}
natsal_final <- update(natsal_sag,
                       m_nonew = ifelse(r_sex == "w", hetnonew, samnonew),
                       w_nonew = ifelse(r_sex == "m", hetnonew, samnonew),
                       r_demo = paste(r_sex, r_sexid, r_sexact, sep = "_"),
                       r_sexid_tog = "all",
                       r_demo_tog = paste(r_sex, r_sexid_tog, r_sexact, sep = "_"))
```

## Analysis

### All sexids, stratified

#### Mean new partner rate

```{r}
all_sexid <- svyby(~m_nonew + w_nonew,
                   by = ~r_sex + r_sexid + r_sexact,
                   design = natsal_final, FUN = svymean)

# for easier use in the model, we give up male and female partnerships their own row
reshape_pt_table <- function(tab) {
  m_only <- select(tab, r_sex, r_sexid, r_sexact, pt_p = m_nonew, pt_p.se = se.m_nonew) %>%
    mutate(rp_sex = "m")
  w_only <- select(tab, r_sex, r_sexid, r_sexact, pt_p = w_nonew, pt_p.se = se.w_nonew) %>%
    mutate(rp_sex = "w")
  tog <- rbind(m_only, w_only) %>%
    arrange(r_sex, r_sexid, r_sexact) %>%
    mutate(r_demo = paste(r_sex, r_sexid, r_sexact, sep = "_")) %>%
    select(r_sex, r_sexid, r_sexact, r_demo, rp_sex, pt_p, pt_p.se)
  return(tog)
}

all_sexid_pt_long <- reshape_pt_table(all_sexid)

all_sexid_pt_long %>%
  format_table()
all_sexid_rep <- all_sexid_pt_long
use_data(all_sexid_rep, overwrite = TRUE)
```

#### Proportions

```{r}
all_sexid_props <- data.frame(svymean(~r_demo, design = natsal_final))
# improve group names
all_sexid_props$r_demo <- str_sub(rownames(all_sexid_props), start = 7L)
rownames(all_sexid_props) <- NULL
split_demo <- function(tab) {
  demos <- str_split(tab$r_demo, "_", simplify = TRUE)
  tab$r_sex <- demos[, 1]
  tab$r_sexid <- demos[, 2]
  tab$r_sexact <- demos[, 3]
  tab
}
all_sexid_props <- all_sexid_props %>%
  split_demo() %>%
  select(r_sex, r_sexid, r_sexact, r_demo, prop = mean, prop_se = SE)

all_sexid_props %>%
  format_table()

use_data(all_sexid_props, overwrite = TRUE)
```


### All sexids, merged together

#### Mean new partner rate

```{r}
sexid_tog <- svyby(~m_nonew + w_nonew, by = ~r_sex + r_sexid_tog + r_sexact,
                   design = natsal_final, FUN = svymean) %>%
  rename(r_sexid = r_sexid_tog)

sexid_tog_long <- reshape_pt_table(sexid_tog)

sexid_tog_long %>%
  format_table()
sexid_tog_rep <- sexid_tog_long
use_data(sexid_tog_rep, overwrite = TRUE)
```

#### Proportions

```{r}
all_sexid_tog_props <- data.frame(svymean(~r_demo_tog, design = natsal_final))
# improve group names
all_sexid_tog_props$r_demo <- str_sub(rownames(all_sexid_tog_props), start = 11L)
rownames(all_sexid_tog_props) <- NULL
all_sexid_tog_props <- all_sexid_tog_props %>%
  split_demo() %>%
  select(r_sex, r_sexid, r_sexact, r_demo, prop = mean, prop_se = SE)

all_sexid_tog_props %>%
  format_table()
sexid_tog_props <- all_sexid_tog_props
use_data(sexid_tog_props, overwrite = TRUE)
```


### Heterosexual only

```{r}
het <- subset(natsal_final, r_sexid == "het")
het_sexid <- svyby(~m_nonew + w_nonew, by = ~r_sex + r_sexid + r_sexact, design = het, svymean)
# assuming 0 between same-sex
het_sexid["m.het.high", "m_nonew"] <- 0
het_sexid["m.het.high", "se.m_nonew"] <- 0
het_sexid["m.het.low", "m_nonew"] <- 0
het_sexid["m.het.low", "se.m_nonew"] <- 0 
het_sexid["w.het.high", "w_nonew"] <- 0
het_sexid["w.het.high", "se.w_nonew"] <- 0 
het_sexid["w.het.low", "w_nonew"] <- 0
het_sexid["w.het.low", "se.w_nonew"] <- 0

# reshape table in a longer format
het_sexid_long <- reshape_pt_table(het_sexid)
het_sexid_long %>%
  format_table()
# rename
het_rep <- het_sexid_long
use_data(het_rep, overwrite = TRUE)
```

#### Proportions

```{r}
het_props <- as.data.frame(svymean(~r_demo, design = het))
# improve group names
het_props$r_demo <- str_sub(rownames(het_props), start = 7L)
rownames(het_props) <- NULL
het_props <- het_props %>%
  split_demo() %>%
  select(r_sex, r_sexid, r_sexact, r_demo, prop = mean, prop_se = SE)

het_props %>%
  format_table()
use_data(het_props, overwrite = TRUE)
```
