# MSID

## Estimate

```{r}
library(reshape2)
msid_est <- all_sex %>%
    group_by(r_sex, r_sexid, r_sexact) %>%
    summarize(het = weighted.mean(hetnonew, w = total_wt),
              hetSD = weighted.sd(hetnonew, w = total_wt),
              sam = weighted.mean(samnonew, w = total_wt),
              samSD = weighted.sd(samnonew, w = total_wt)) %>%
    mutate(m = ifelse(r_sex == "m", sam, het),
           msd = ifelse(r_sex == "m", samSD, hetSD),
           w = ifelse(r_sex == "w", sam, het),
           wsd = ifelse(r_sex == "w", samSD, hetSD)) %>%
    select(r_sex, r_sexid, r_sexact, m, msd, w, wsd)

msid_rep <- msid_est %>%
    select(r_sex, r_sexid, r_sexact, m, w) %>%
    melt(id.vars = c('r_sex', 'r_sexid', 'r_sexact'),
         variable.name = "rp_sex",
         value.name = "partners") %>%
    # add proportions
    rowwise() %>%
    mutate(prop = get_proportion(all_sex, q_sex = r_sex,
                                q_sexid = r_sexid, q_sexact = r_sexact),
           n_partners = prop * partners) %>%
    ungroup() %>%
  arrange(r_sex, r_sexid, r_sexact)
format_table(msid_rep)
```


```{r}
msid_offered <- msid_rep %>%
  group_by(rp_sex, r_sex) %>%
  # define prop_to_r as the the proportion of all partners offered to r that come from rp
  mutate(prop_1to2 = n_partners / sum(n_partners)) %>%
  select(sex1 = r_sex,
         sexid1 = r_sexid,
         sexact1 = r_sexact,
         sex2 = rp_sex,
         prop_1to2) %>%
  arrange(sex1, sexid1, sexact1, sex2)
format_table(msid_offered)
```

Note that most of the partners that are 'offered' from men to men from men come from high sexual-activity gay men, and the proportion of all partnerships offered from men to men sums to 1.

```{r}
fmtm <- msid_offered %>%
  filter(sex1 == "m" & sex2 == "m") %>%
  mutate(rp_demo = paste(sex1, sexid1, sexact1, sep="_"))
ggplot(fmtm) +
  geom_bar(aes(x = rp_demo, y = prop_1to2), stat = "identity") +
  ylab("Proportion of partnerships offered to men from men") +
  theme_bw()
```

Make partner distribution

## Distribute

```{r}
make_partner_dist_msid <- function(report, dist){
    left_join(report, dist,
          by = c("rp_sex" = "sex1", "r_sex" = "sex2")) %>%
        mutate(d_part_p = partners * prop_1to2) %>%
        mutate(r_demo = paste(r_sex, r_sexid, r_sexact, sep="_"),
               rp_demo = paste(rp_sex, sexid1, sexact1, sep="_")) %>%
  select(r_demo, r_sex, r_sexid, r_sexact,
         rp_demo, rp_sex, rp_sexid = sexid1, rp_sexact = sexact1, prop, d_part_p)
}

msid_partner_dist <- make_partner_dist_msid(msid_rep, msid_offered)

format_table(head(msid_partner_dist, n = 10))
```

Look at bi men
```{r}
men_bi_high <- msid_partner_dist %>%
  filter(r_demo == "m_bi_high" & str_detect(rp_demo, "w_"))
format_table(men_bi_high)
sum(men_bi_high$d_part_p)
```


## Balance

```{r}
msid_bidi <- make_bidirectional(msid_partner_dist)
contact_msid_base <- balance_het(msid_bidi)
format_table(head(contact_msid_base))
```

Check that this worked:

```{r}
all(with(contact_msid_base, abs(cnr - cnrp) < .Machine$double.eps), na.rm = TRUE)
```

### Compare
```{r}
sexid_nos_by_grp <- contact_msid_base %>%
  mutate(rp_sex = str_split(rp_demo, "_", simplify = TRUE)[, 1]) %>%
  group_by(r_demo, rp_sex) %>%
  summarize(bal_p_pc = sum(corrected_r, na.rm = TRUE))

sexid_nos_compare <- msid_rep %>%
  mutate(r_demo = paste(r_sex, r_sexid, r_sexact, sep = "_")) %>%
  select(r_demo, rp_sex, orig_p_pc = partners) %>%
  left_join(sexid_nos_by_grp, by = c("r_demo", "rp_sex"))
format_table(sexid_nos_compare)
```

### Plot and save data

```{r}
ggplot(rbind(contact_msid_base)) +
    geom_tile(aes(x = r_demo, y = rp_demo, fill=corrected_r), color="black", size = 0.2) +
    scale_fill_gradient2(name="Annual Number\nof New Partnerships",
                         low = "white",
                         mid = "blue2",
                         high = "black",
                         midpoint = 3,
                         na.value = "white") +
    labs(x = "Group of Partner 1",
         y = "Group of Partner 2") +
    theme_minimal(base_size = 14) + coord_fixed() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

# save plot and data
ggsave("plots/msid_partner_dist.png", width = 7, height=6, units = 'in', dpi =500)

## columns we need
contact_msid_base <- contact_msid_base %>%
  select(r_sex, r_sexid, r_sexact, r_demo,
         rp_sex, rp_sexid, rp_sexact, rp_demo, prop = prop.r, corrected_r)
usethis::use_data(contact_msid_base, overwrite = TRUE)
```

## Define prior distributions for calibration

Calculate gamma params

### Partnerships with Men
```{r fig.height = 10}
# men
msid_m_gam_params <- lapply(1:12,
                          function(i) {
                              p <- gamma_params(msid_est$m[i],
                                           msid_est$msd[i])
                              p$name <- paste0(with(msid_est, paste(r_sex[i], r_sexid[i], r_sexact[i], sep = "_")))
                              return(p)
                              })
par(mfrow = c(4, 3))
devnull <- lapply(msid_m_gam_params,
       function(p) {
           if (is.nan(p$shape)) {
               return(NULL)
           } else {
               curve(dgamma(x, shape = p$shape, rate = p$rate),
                 from = 0, to = p$shape / p$rate + 2 * p$shape / p$rate^2,
                 main = p$name,
                 ylab = "Density")
           }
           }
       )
```

### Partnerships with Women
```{r fig.height = 10}
msid_w_gam_params <- lapply(1:12,
                          function(i) {
                              p <- gamma_params(msid_est$w[i],
                                           msid_est$wsd[i])
                              p$name <- paste0(with(msid_est, paste(r_sex[i], r_sexid[i], r_sexact[i], sep = "_")))
                              return(p)
                              })
par(mfrow = c(4, 3))
devnull <- lapply(msid_w_gam_params,
       function(p) {
           if (is.nan(p$shape)) {
               return(NULL)
           } else {
               curve(dgamma(x, shape = p$shape, rate = p$rate),
                 from = 0, to = p$shape / p$rate + 2 * p$shape / p$rate^2,
                 main = p$name,
                 ylab = "Density")
           }
           }
       )
```
